---
title: "Human Activity Recognition using Random Forest"
author: "Hang Su"
date: "December 27, 2015"
output: html_document
---

In this project, We will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict their exercise manner. [Traing](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) and [test](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv) data is downloaded from [this website](http://groupware.les.inf.puc-rio.br/ha).

# Load data
```{r}
library(randomForest)
traindata <- read.csv("pml-training.csv")
testdata <- read.csv("pml-testing.csv")
head(traindata[,15])
```
Load the data and take a look at it.

# Building model
As is shown before, many columns are empty or ```NA```. After checking all the columns, we select following features: ```{roll, pitch, yaw, total_accel} x {belt, arm, dumbbell, forearm}```, i.e. 16 in total.

We Perform K-fold cross validation with k = 5.
```{r}
accuracy <- function(model,testset){sum(predict(fit, testset) == testset$classe)/nrow(testset)}

set.seed(7)
order <- sample(nrow(traindata), size = nrow(traindata), replace = FALSE)

k <- 5
nrowspercv <- nrow(traindata) / k
for (i in 1:k){
  part <- order[((i-1)*nrowspercv+1):min(i*nrowspercv, nrow(traindata))]
  
  fit <- randomForest(classe ~ roll_belt + pitch_belt + yaw_belt + total_accel_belt + roll_arm + pitch_arm + yaw_arm + total_accel_arm + roll_dumbbell + pitch_dumbbell + yaw_dumbbell + total_accel_dumbbell + roll_forearm + pitch_forearm + yaw_forearm + total_accel_forearm, method = "rf", data = traindata[-part,], nodesize=5, allowParallel=TRUE)
  acc <- accuracy(fit, traindata[part,])
  print(paste("cv", i, "acc:", acc * 100, "%"))
}
```

# Out of sample error analysis
As is shown in the previous section, accuracies on cross validation set are higher than 98%.

We projection data points to 2-d space, and plot the location.
```{r}
library(ggplot2)
qplot(total_accel_forearm, pitch_forearm, col = classe, data = traindata)
```

We could also analyze the features using ```featurePlot```.
```{r}
part <- order[1:nrowspercv]
library(caret)
featurePlot(x=traindata[part,c("roll_belt", "pitch_forearm", "total_accel_dumbbell", "yaw_forearm")], y=traindata$classe[part], plot = "pairs")
```

The expected out of sample error could be due to noises in sensors or differences between participants.

# Predict output for test set
Now, train a model with all the training data and predict for test data
```{r}
fit <- randomForest(classe ~ roll_belt + pitch_belt + yaw_belt + total_accel_belt + roll_arm + pitch_arm + yaw_arm + total_accel_arm + roll_dumbbell + pitch_dumbbell + yaw_dumbbell + total_accel_dumbbell + roll_forearm + pitch_forearm + yaw_forearm + total_accel_forearm, method = "rf", data = traindata, nodesize=5, allowParallel=TRUE)
answers <- predict(fit, testdata)
```
Take a look at the answer
```{r}
answers
```